---
title: "Seurat - Male and female germ cells combined"
output:
  html_document:
    keep_md: true
    smart: false
    theme: united
date: 'Compiled: `r format(Sys.Date(), "%B %d, %Y")`'
---
***


### Human Spermatogenesis and female Fetal Germ Cells
* Data from Wang et al (2018) and L. Li et all.(DOI: 10.1016/j.stem.2017.03.007)
```{r, message=FALSE}
## Libraries
suppressMessages(library(gridExtra))
suppressMessages(library(ggplot2))
suppressMessages(library(Seurat))
suppressMessages(library(ggplot2))
suppressMessages(library(plotly))
suppressMessages(library(dplyr))
suppressMessages(library(genefilter))
suppressMessages(library(gplots))

outputDir  = getwd()
```

### Load data
```{r, message=FALSE}
## Load male data, select the clusters of interest
male <- readRDS("/path/to/248-FGC_RNAseq/analysis/maleSeurat/revisit/male_mnn.rds")
maleMeiosis <- subset(male, ident=c(0,1,3,5))
maleMeiosis <- RenameIdents(object = maleMeiosis, 
                            '0' = 'M_Z-eP',
                            '1' = 'M_L', 
                            '3' = 'M_P',
                            '5' = 'M_D')
maleMeiosis$original_ids <- Idents(maleMeiosis)

## Load FGCs, late meiotic
lateMeiot <- readRDS("/path/to/248-FGC_RNAseq/analysis/LiLiSeurat/revisitLateMeiot/lateMeiot.rds")
lateMeiot <- subset(lateMeiot, ident=c(1,2,3,4))
lateMeiot <- RenameIdents(object = lateMeiot,
                          '1' = 'F_L',
                          '4' = 'F_Z-eP',
                          '3' = 'F_P',
                          '2' = 'F_D')
lateMeiot$original_ids <- Idents(lateMeiot)

## Combine datasets
maleFemaleMeiosis <- merge(maleMeiosis, y=lateMeiot, add.cell.ids = c("M_", "F_"))
## Cells in each dataset
dim(maleMeiosis)
dim(lateMeiot)
```

### Confirm the saved male dataset is as expected by ploting the tSNE
```{r}
 DimPlot(object = male, 
             reduction = "tsne")
```

## Share of cluster IDs in male cells
```{r}
cluster_table <- table(Idents(maleMeiosis))
cluster_table
cluster_table/sum(cluster_table)
```

### Fix numerical imbalance between datasets
* There are 1312 cells in the male meiotic cells dataset, as oposed to 158 in the female. It is probably better for the male/female comparison to keep the numbers on par. Subsample the male to get a similar number of cells (160)
* We want to have the per cluster share of males cells comparable before and after sampling.  set.seed(11) gives satisfactory results
* We want the male cell selection to be reproducible. We save once the cell IDs in a file and use them for any future workflow run
```{r, message=FALSE}
set.seed(11)
sampledCellIDs <- sample(x = colnames(maleMeiosis), size = 160, replace = FALSE)
maleMeiosisSubsmpl <- subset(x = maleMeiosis, cells=sampledCellIDs) 
cluster_table <- table(Idents(maleMeiosisSubsmpl))
cluster_table
cluster_table/sum(cluster_table)
maleFemaleMeiosisEqualized <- merge(maleMeiosisSubsmpl, y=lateMeiot, add.cell.ids = c("M_", "F_"))
```

# Resume with the analysis workflow 
## Process the combination with meiotic male intact
### Find Variable Genes
```{r, message=FALSE}
maleFemaleMeiosis <- FindVariableFeatures(object = maleFemaleMeiosis, 
                         mean.function = ExpMean, 
                         dispersion.function = LogVMR,
                         nfeatures = 2500)
length(x = VariableFeatures(maleFemaleMeiosis))
topAveExpr = HVFInfo(maleFemaleMeiosis)[HVFInfo(maleFemaleMeiosis)[,1]>3, ]
topAveExprPath = paste0(outputDir, "/topAveExpr.tsv")
write.table(x = topAveExpr, file = topAveExprPath, sep = "\t")
```

### Scaling the data
```{r, message=FALSE}
maleFemaleMeiosis <- ScaleData(object = maleFemaleMeiosis)
```

### PCA
```{r, message=FALSE}
maleFemaleMeiosis <- RunPCA(object = maleFemaleMeiosis, 
              features = VariableFeatures(maleFemaleMeiosis), 
              do.print = TRUE, 
              pcs.print = 1:5, 
              genes.print = 5)
```

### Vizualize PCA loadings
```{r, message=FALSE}
VizDimLoadings(object = maleFemaleMeiosis, 
               dims = 1:2, 
               reduction = "pca")
ggsave(paste0(outputDir, "/vizPCA.pdf"), width = 10, height = 7)
DimPlot(object = maleFemaleMeiosis, reduction = "pca")
write.table(Loadings(object = maleFemaleMeiosis, reduction = "pca"), file = paste0(outputDir, "/geneLoadings.tsv"), sep = "\t", quote = FALSE)
```

### Elbow plot
```{r, message=FALSE}
ElbowPlot(object = maleFemaleMeiosis)
## Print the percenatge of the standard deviation of each PC as a 
## fraction of the total standard deviation od the first 20 PCs
totalSdev <- sum(Stdev(object = maleFemaleMeiosis, reduction = "pca"))
print(Stdev(object = maleFemaleMeiosis, reduction = "pca")/totalSdev)
```

### Calculate clusters, tSNEs and UMAP
```{r, message=FALSE}
maleFemaleMeiosis <- FindNeighbors(maleFemaleMeiosis,
                                   reduction = "pca",
                                   dims = 1:11)
maleFemaleMeiosis <- FindClusters(object = maleFemaleMeiosis, 
                                  resolution = 0.6)
```

```{r, message=FALSE}
maleFemaleMeiosis <- RunTSNE(object = maleFemaleMeiosis, 
                             dims = 1:11)
maleFemaleMeiosis <- RunUMAP(object = maleFemaleMeiosis, 
                             dims = 1:11)
```

### Plot tSNEs UMAP
```{r, message=FALSE}
DimPlot(object = maleFemaleMeiosis, 
        reduction = "tsne")
ggsave(paste0(outputDir, "/tSNE.pdf"), width = 10, height = 7)
p <- DimPlot(object = maleFemaleMeiosis, 
             reduction = "umap")
ggplotly(p, width = 900, height = 700)
ggsave(paste0(outputDir, "/UMAP.pdf"), width = 10, height = 7)
DimPlot(object = maleFemaleMeiosis, 
        reduction = "pca")
ggsave(paste0(outputDir, "/PCA.pdf"), width = 10, height = 7)
```

## Get gender of cells and plot
```{r, message=FALSE}
gender <- sapply(colnames(x = maleFemaleMeiosis), function(x){
  strsplit(x, split = "__", fixed = TRUE)[[1]][1]
})
maleFemaleMeiosis$gender <- gender
p <- DimPlot(object = maleFemaleMeiosis, 
             reduction = "tsne",
             group.by = "gender")
ggsave(paste0(outputDir, "/tSNE_gender.pdf"), width = 10, height = 7)
```

===================================================

## Process the combination with male meitoic subsampled
### Plot the selected cells on the plot with all male cells
Check if the subsampled cells are representative of the male cell population
```{r, message=FALSE}
## Add the M_ in the cell ID
sampledIDs <- paste0("M__",sampledCellIDs)
p <- DimPlot(object = maleFemaleMeiosis, 
             reduction = "tsne",
             cells.highlight = sampledIDs)
ggplotly(p, width = 900, height = 700)
ggsave(paste0(outputDir, "/tSNE_sampled.pdf"), width = 10, height = 7)
```

# Work with the male female equalized dataset
### Find Variable Genes
```{r, message=FALSE}
maleFemaleMeiosisEqualized <- FindVariableFeatures(object = maleFemaleMeiosisEqualized, 
                         mean.function = ExpMean, 
                         dispersion.function = LogVMR,
                         nfeatures = 2500)
length(x = VariableFeatures(maleFemaleMeiosisEqualized))
topAveExpr = HVFInfo(maleFemaleMeiosisEqualized)[HVFInfo(maleFemaleMeiosisEqualized)[,1]>3, ]
topAveExprPath = paste0(outputDir, "/topAveExpr.tsv")
write.table(x = topAveExpr, file = topAveExprPath, sep = "\t")
```

### Scaling the data
```{r, message=FALSE}
maleFemaleMeiosisEqualized <- ScaleData(object = maleFemaleMeiosisEqualized)
```

### PCA
```{r, message=FALSE}
maleFemaleMeiosisEqualized <- RunPCA(object = maleFemaleMeiosisEqualized, 
                                     features = VariableFeatures(maleFemaleMeiosisEqualized), 
                                     do.print = TRUE, 
                                     pcs.print = 1:5, 
                                     genes.print = 5)
```

### Vizualize PCA loadings
```{r, message=FALSE}
VizDimLoadings(object = maleFemaleMeiosisEqualized, 
               dims = 1:2, 
               reduction = "pca")
ggsave(paste0(outputDir, "/vizPCA_Eq.pdf"), width = 10, height = 7)
DimPlot(object = maleFemaleMeiosisEqualized, reduction = "pca")
write.table(Loadings(object = maleFemaleMeiosisEqualized, reduction = "pca"), file = paste0(outputDir, "/geneLoadings.tsv"), sep = "\t", quote = FALSE)
```

### Elbow plot
```{r, message=FALSE}
ElbowPlot(object = maleFemaleMeiosisEqualized)
## Print the percenatge of the standard deviation of each PC as a 
## fraction of the total standard deviation od the first 20 PCs
totalSdev <- sum(Stdev(object = maleFemaleMeiosisEqualized, reduction = "pca"))
print(Stdev(object = maleFemaleMeiosisEqualized, reduction = "pca")/totalSdev)
```

### tSNEs and UMAP
```{r, message=FALSE}
maleFemaleMeiosisEqualized <- RunTSNE(object = maleFemaleMeiosisEqualized, 
                             dims = 1:11)
maleFemaleMeiosisEqualized <- RunUMAP(object = maleFemaleMeiosisEqualized, 
                             dims = 1:11)
```

### Plot tSNEs UMAP
```{r, message=FALSE}
DimPlot(object = maleFemaleMeiosisEqualized, 
        reduction = "tsne",
        group.by = "original_ids")
ggsave(paste0(outputDir, "/tSNE_Eq.pdf"), width = 10, height = 7)
DimPlot(object = maleFemaleMeiosisEqualized, 
        reduction = "umap",
        group.by = "original_ids")
# ggplotly(p, width = 900, height = 700)
ggsave(paste0(outputDir, "/UMAP_Eq.pdf"), width = 10, height = 7)
DimPlot(object = maleFemaleMeiosisEqualized, 
        reduction = "pca",
        group.by = "original_ids")
ggsave(paste0(outputDir, "/PCA_Eq.pdf"), width = 10, height = 7)
```

### Save genderof cells and plot
```{r, message=FALSE}
gender <- sapply(colnames(x = maleFemaleMeiosisEqualized), function(x){
  strsplit(x, split = "__", fixed = TRUE)[[1]][1]
})
maleFemaleMeiosisEqualized$gender <- gender
DimPlot(object = maleFemaleMeiosisEqualized, 
        reduction = "umap",
        group.by = "gender")
ggsave(paste0(outputDir, "/umap_Eq_gender.pdf"), width = 10, height = 7)
```

### Calculate clusters
```{r, message=FALSE}
maleFemaleMeiosisEqualized <- FindNeighbors(maleFemaleMeiosisEqualized,
                                   reduction = "pca",
                                   dims = 1:11)
maleFemaleMeiosisEqualized <- FindClusters(object = maleFemaleMeiosisEqualized, 
                                  resolution = 1)
```

### Plot tSNEs UMAP again, with the newly calculated clusters
```{r, message=FALSE}
DimPlot(object = maleFemaleMeiosisEqualized, 
        reduction = "tsne")
ggsave(paste0(outputDir, "/tSNE_Eq_recalc.pdf"), width = 10, height = 7)
p <- DimPlot(object = maleFemaleMeiosisEqualized, 
             reduction = "umap")
ggplotly(p, width = 900, height = 700)
ggsave(paste0(outputDir, "/UMAP_Eq_recalc.pdf"), width = 10, height = 7)
DimPlot(object = maleFemaleMeiosisEqualized, 
        reduction = "pca")
ggsave(paste0(outputDir, "/PCA_Eq_recalc.pdf"), width = 10, height = 7)
```

### Plot gender
```{r, message=FALSE}
DimPlot(object = maleFemaleMeiosisEqualized, 
        reduction = "umap",
        group.by = "gender")
ggsave(paste0(outputDir, "/umap_Eq_gender_recalc.pdf"), width = 10, height = 7)
saveRDS(maleFemaleMeiosisEqualized, file = paste0(outputDir, "/maleFemaleMeiosisEqualized.rds"))
```


### SessionInfo
```{r}
sessionInfo()
```
